Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-09-28 packages/spec-haskell/yellowpaper/Paper.tex
section Relevant Approaches  done!

\begin{itemize}
\item U is the set of monetary units.
\item $\nu : U \rightarrow \mathbb{N}$ is the value function defining the value $\nu(u)$ of every value unit u. The set
    $\mathbb{N}$ is the set of all natural numbers, but instead, we can use any set of numerals that is totally ordered
    (\eg\ integers, real numbers).
\item $\beta : U \rightarrow \mathfrak{B}$ is the bearer function defining the bearer $\beta(u)$ of a unit. The set
    $\mathfrak{B}$ is the set of possible bearers. The bearer is usually a legal construction defining any type of legal
    entity, such as a person, a family, a company, a state institution, etc.
\end{itemize}

@@

Recall that we want our modernized payment system to handle money distribution continuously over time, and its financial
contracts compositional. A very closely related software design paradigm best known to address those needs is
\textit{functional reactive programming (FRP)}. It was first introduced by Conal Elliott \& Paul Hudak in solving
multimedia animations \cite{elliott1997functional}. Later on, Hudak also worked on \cite{hudak2002arrows} and
\cite{wan2000functional} making FRP a more general framework for programming hybrid system with continuous behaviors in
a high-level, declarative manner.

@@

Values that vary over continuous time are called \textit{behaviors}. They are first-class values, and are built up
compositionally. In modern payment systems, we should want the monetary values of monetary units capable of varying over
continuous time. The semantic function of $\alpha$-behaviors produces the value of type $\alpha$ of a behavior at a given
time:

@@

In modern payment systems, the payment primitives executed in payment execution environments are one type of
events. More types of events can be read from the original paper \footfullcite[][section 2.3 Semantics of
    Events]{elliott1997functional}.

@@

In the context of modern payment systems, the occurrences of payment primitives (events) change the behaviors of
monetary units in how much monetary values it has over continuous. The building blocks of the financial contracts should
be about modeling these events and their reactivity declaratively.

@@

We also want a formal and precise specification of payment primitives. The title of the paper contains
\textit{denotational semantics}: ``it is a \textit{compositional} style for precisely specifying the meanings of
languages, invented by Christopher Strachey and Dana Scott in the 1960s (\cite{scott1971toward})'', and Conal Elliott
proposed that denotational semantics can also be applied to \textit{data types} in programming \footfullcite[][section
    2, denotational semantics and data types]{Elliott2009-type-class-morphisms-TR}. We will revisit this formulation
too.

... TODO

>>>>

\begin{itemize}
\item U is the set of monetary units.
\item $\nu : U \rightarrow \mathbb{N}$ is the value function defining the value $\nu(u)$ of every value unit u. The set
    $\mathbb{N}$ is the set of all natural numbers, but instead, we can use any set of numerals that is totally ordered
    (\eg\ integers, real numbers).
\item $\beta : U \rightarrow \mathcal{B}$ is the bearer function defining the bearer $\beta(u)$ of a unit. The set
    $\mathcal{B}$ is the set of possible bearers. The bearer is usually a legal construction defining any type of legal
    entity, such as a person, a family, a company, a state institution, etc.
\end{itemize}

@@

Recall that we want our modernized payment systems to handle money distribution continuously over time, and its financial
contracts compositional. A very closely related software design paradigm best known to address those needs is
\textit{functional reactive programming (FRP)}. It was first introduced by Conal Elliott \& Paul Hudak in solving
multimedia animations \cite{elliott1997functional}. Later on, Hudak also worked on \cite{hudak2002arrows} and
\cite{wan2000functional} making FRP a more general framework for programming hybrid system with continuous behaviors in
a high-level, declarative manner.

@@

Values that vary over continuous time are called \textit{behaviors}. They are first-class values, and are built up
compositionally. In modernized payment systems, we should want the monetary values of monetary units capable of varying over
continuous time. The semantic function of $\alpha$-behaviors produces the value of type $\alpha$ of a behavior at a given
time:

@@

In modernized payment systems, the payment primitives executed in payment execution environments are one type of
events. More types of events can be read from the original paper \footfullcite[][section 2.3 Semantics of
    Events]{elliott1997functional}.

@@

In the context of modernized payment systems, the occurrences of payment primitives (events) change the behaviors of
monetary units in how much monetary values it has over continuous. The building blocks of the financial contracts should
be about modeling these events and their reactivity declaratively.

@@

We also want a formal and precise specification of payment primitives. The title of the paper includes
\textit{denotational semantics}: ``it is a \textit{compositional} style for precisely specifying the meanings of
languages, invented by Christopher Strachey and Dana Scott in the 1960s (\cite{scott1971toward})'', and Conal Elliott
proposed that denotational semantics can also be applied to \textit{data types within a programming language}
\footfullcite[][section 2, denotational semantics and data types]{Elliott2009-type-class-morphisms-TR}.

To create denotational semantics for each syntactic category $\mathcal{C}$, we should specify:

\begin{itemize}
\item a mathematical model $[\![\mathcal{C}]\!]$ of meanings, and
\item a semantic function $[\![.]\!]_{\mathcal{C}} :: \mathcal{C} \rightarrow [\![\mathcal{C}]\!]$.
\end{itemize}

The syntactic category we are interested in is \textit{payment primitives}, which we should treat them as FRP-style
\textit{Behavior} data types. In the following chapters, it is also unambiguously referred to in short as
$[\![.]\!]$. Various $[\![.]\!]$ must be compositional, \ie must be defined by structural recursion.

It is important to note that our purpose of using the denotational semantics is to give precise meaning to payment
primitives, independent of its implementation (which deals with performance, optimization, side effects, etc.). We will
spend the next chapter exploring the denotational semantics for general payment primitives in modernized payment
systems.
